### 2014-01-22 presentations/popl-2013/minute-madness.tex
Fix some broken tex

\documentclass[utf8x]{beamer}

\usepackage{beamerthemesplit}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}
\usepackage[all]{xy}

\pdfinfo {
  /Title (POPL --- Minute Madness)
  /Subject (POPL)
  /Author (Jason Gross)
  /Keywords (POPL, Coq, PLMW, Minute Madness, Category Theory)
}

\begin{document}

\title{POPL --- Minute Madness}
\date{POPL 2013}
\author{Jason Gross --- \href{mailto:jgross@mit.edu}{jgross@mit.edu}}

\begin{frame}{Database Management on top of Category Theory in Coq \\ Category of Relational Schemas $\cong$ Category of Categories}
  \begin{center}
    \fbox{
      $$
        \xymatrix{
          a \ar@/^{0.5pc}/[rr]^{f} \ar@/_{0.5pc}/[d]_{g} \ar@/^{0.5pc}/[d]^{h} && b \ar@/^{0.5pc}/[ll]^{i} \ar@/^{1.5pc}/[dll]^{j} \\
          c
        }
        $$
      } \raisebox{-3.5ex}[\height][\depth]{$\cong$}
    \fbox{\small
      $$
        \xymatrix{
          *\txt{professor} \ar@/^{1pc}/[rr]^{\text{department}} \ar@/_{1pc}/[dd]_{\text{f.~name}} \ar@/^{1pc}/[dd]^{\text{l.~name}} && *\txt{department} \ar@/^{1pc}/[ll]^{\text{department head}} \ar@/^{2pc}/[ddll]^{\text{name}} \\ \\
          *\txt{string} &&
        }
        $$
      }

    \vspace{6ex} The diagrams are ``the same''.
  \end{center}
\end{frame}
\end{document}