### 2021-08-15 resume/Resume.tex
Revert "Deploying to gh-pages from @ JasonGross/JasonGross.github.io@cbda5049e77b1054b88d69e586a21a32885f0aa8 🚀"
This reverts commit 8d3ef534c8997c7ef4e469dcf58ccdd69af0346b.

% LaTeX file for resume
% This file uses the resume document class (res.cls)

\documentclass[11pt]{res}
\usepackage[T1]{fontenc}
\usepackage{xifthen,verbatim}
\usepackage{hyperref}
\usepackage{etoolbox}[2011/01/03]

\usepackage[backend=biber,bibencoding=utf-8,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}

\makeatletter

\let\origifempty\ifempty
\undef\ifempty
\let\origforlistloop\forlistloop
\usepackage{etextools}
\let\eifempty\ifempty
\let\ifempty\origifempty
\let\forlistloop\origforlistloop
\def\preetextools{\let\ifempty\eifempty}

\newif\ifmultitexcompile
\multitexcompiletrue

\newcommand\multitex@separator{-}
\newwrite\multitex@temp
% \immediate\openout\multitex@temp{\jobname.mtx}
\newtoks\multitex@names
\multitex@names={}
\newcommand{\multitex@postprocess}[1]{%
  \immediate\expandafter\closeout\csname @multitex@#1@file\endcsname
  \ifmultitexcompile
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
    \immediate\write18{biber \csname @multitex@#1@barename\endcsname}%
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
  \fi
}
\newcommand{\onlymultitex}[1]{\FE@ifstar{#1}{\only@other@multitex}{\only@multitex}}
\newcommand{\only@multitex}[2]{#2}
\newcommand{\only@other@multitex}[2]{}
{\catcode`\#=12
  \global\def\@pound{#}
}
\newcommand{\DeclareFile}[2][]{
  \global\multitex@names=\expandafter{\the\multitex@names,#2}
  \expandafter\xdef\csname @multitex@#2\endcsname{\xifempty{#1}{#2}{#1}}
  \expandafter\xdef\csname @multitex@#2@name\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.tex"}
  \expandafter\xdef\csname @multitex@#2@auxname\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.aux"}
  \expandafter\xdef\csname @multitex@#2@barename\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname"}
  \expandafter\newwrite\csname @multitex@#2@file\endcsname
  \immediate\expandafter\openout\csname @multitex@#2@file\expandafter\endcsname\csname @multitex@#2@name\endcsname
  \AtBeginDocument{%
    \multitex@write{#2}{%
      \detokenize{\newcommand{\onlymultitex}[1]}%
        {\detokenize{\csname FE@ifstar\endcsname}%
          {\@pound1}%
          \detokenize{%
            {\csname only@multitex\endcsname}{\csname only@multitex\endcsname}%
          }%
        }%
      \detokenize{%
        \expandafter\newcommand\csname only@multitex\endcsname[2]}%
        {%
          \noexpand\renewcommand*{\noexpand\do}[1]{%
            \noexpand\ifstrequal{\@pound\@pound1}{#2}%
              {\@pound2\noexpand\listbreak}%
              {}%
            }%
          {\noexpand\edef\noexpand\foo{\@pound1}\noexpand\expandafter}%
          \noexpand\expandafter\noexpand\docsvlist\noexpand\expandafter{\noexpand\foo}%
        }
      \detokenize{%
        \expandafter\let\csname only@other@multitex\expandafter\endcsname\csname only@multitex\endcsname
      }%
    }%
  }
  \AtEndDocument{\multitex@postprocess{#2}}
}
\newcommand{\multitex@write}[2]{\xifblank{#1}{}{\immediate\expandafter\write\csname @multitex@#1@file\endcsname{#2}}}
\edef\multitex@tempfilename{\jobname.mtx}
\newcommand{\multitex@pre@writetothis}{\immediate\openout\multitex@temp\multitex@tempfilename}%\long\gdef\multitex@thiscode{}}%
\newcommand{\multitex@writetothis}[1]{\immediate\write\multitex@temp{#1}}%\ExpandNext{\long\gappto\multitex@thiscode}{\expandnext{\relax\scantokens}{#1}}}%
\newcommand{\multitex@post@writetothis}{\immediate\closeout\multitex@temp\relax\input{\multitex@tempfilename}}%\show\multitex@thiscode\multitex@thiscode\relax}%
\newcommand{\multitex@writetoothers}[2][\the\multitex@names]{{\preetextools\ExpandNext{\forcsvloop}{#1}\do{\multitex@write{##1}{#2}}}}
\newcommand{\DeclareDocumentClass}[1]{\multitex@writetoothers{\detokenize{#1}}}
\AtBeginDocument{\multitex@writetoothers{\detokenize{\begin{document}}}}
\AtEndDocument{\multitex@writetoothers{\detokenize{\end{document}}}}
\newcommand{\includein}[2][]{\multitex@pre@writetothis\multitex@writetothis{\detokenize{#2}}\multitex@post@writetothis\multitex@writetoothers[#1]{\detokenize{#2}}}
\AfterEndEnvironment{multitex}{\multitex@post@writetothis}
\newenvironment{multitex}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \multitex@pre@writetothis
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetothis{\the\verbatim@line}\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
% only others
\newenvironment{multitex*}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
\makeatother

\DeclareFile{curriculum-vitae}
\DeclareFile{mathematics}
\DeclareFile{computer-science}
\DeclareFile{curriculum-vitae-long}
\DeclareFile{curriculum-vitae-mathematics}
\DeclareFile{curriculum-vitae-computer-science}
\DeclareDocumentClass{\documentclass[11pt]{res}}

\begin{multitex}
  \edef\resumefiles{mathematics,computer-science}
  \edef\cvshortspecialfiles{curriculum-vitae-mathematics,curriculum-vitae-computer-science}
  \edef\cvshortfiles{curriculum-vitae,\cvshortspecialfiles}
  \edef\cvlongfiles{curriculum-vitae-long}
  \edef\cvfiles{\cvshortfiles,\cvlongfiles}
  \edef\highschoolfiles{curriculum-vitae-long}
  \edef\mathfiles{mathematics,curriculum-vitae-mathematics}
  \edef\cvlongandmathfiles{\mathfiles,curriculum-vitae-long}
  \edef\cvandmathfiles{\cvlongandmathfiles,curriculum-vitae,curriculum-vitae-computer-science}
  \edef\computersciencefiles{computer-science,curriculum-vitae-computer-science}
  \edef\cvlongandcomputersciencefiles{\computersciencefiles,curriculum-vitae-long}
  \edef\cvandcomputersciencefiles{curriculum-vitae,curriculum-vitae-mathematics,\cvlongandcomputersciencefiles}
  \edef\allexceptcv{\resumefiles,\cvshortspecialfiles}
  \edef\allexceptcvlongfiles{\resumefiles,\cvshortfiles}
  \edef\allexceptcvandcvlongfiles{\resumefiles,\cvshortspecialfiles}
  %\usepackage[utf8x]{inputenc}
  %\usepackage[T1]{fontenc}
  %\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
  %\usepackage{newcent}   % uses new century schoolbook postscript font
  \usepackage[T1]{fontenc}
  \usepackage[margin=1in,top=0.5in,bottom=0.5in]{geometry}
  \usepackage{hyperref}
  \usepackage{calc}
  \usepackage{amsmath}
  \usepackage{etoolbox}[2011/01/03]
  %\bibliographystyle{plainyr-rev}
  \usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}
  \addbibresource{jason-gross.bib}
  \DeclareBibliographyCategory{exclude}
  \makeatletter
  \@ifpackageloaded{etextools}{%
  }{%
    \let\origifempty\ifempty
    \let\origforlistloop\forlistloop
    \undef\ifempty
    \usepackage{etextools}
    \let\eifempty\ifempty
    \let\ifempty\origifempty
    \let\forlistloop\origforlistloop
    \def\preetextools{\let\ifempty\eifempty}
  }%
  \makeatother
  \begin{format}
    \employer{l}\dates{r}\\
    \title{l}\location{r} \\
    \body
  \end{format}
  \let\oldemployer=\employer
  \let\oldtitle=\title
  \renewcommand{\employer}[1]{\oldemployer{\textbf{#1}}}
  \renewcommand{\title}[1]{\oldtitle{\textit{#1}}}

  % \let\oldposition=\position
  % \let\oldendposition=\endposition
  % \renewenvironment{position}{\oldposition\vspace*{-2\baselineskip}}{\endoldposition}

  \makeatletter
  \let\old@itemize=\itemize
  \def\itemize{\old@itemize\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}\setlength{\leftskip}{-1em}}
  \renewenvironment{position}
    {%
    \begingroup
      \par
        \the\tabular@head
  %     \addpenalty{-\@secpenalty}% bad place for a page break
      \penalty -\@secpenalty % bad place for a page break
      \penalty 10000
      \ignorespaces
      \vspace*{-\baselineskip}
    }{%
        \the\tabular@tail
  %     \addpenalty{\@secpenalty}% good place for a page break
      \penalty \@secpenalty % good place for a page break
      \endgroup
  }
  \makeatother

  \newsectionwidth{1em}
  \resumewidth=7.5in

  \newcommand{\aswidthof}[2]{\rlap{#1}\hphantom{#2}}
\end{multitex}
\begin{multitex*}[\cvlongfiles]
  \name{\LARGE \bf \hspace{\stretch{1}} Jason Gross \hspace{\stretch{1}} \\
    \hspace{\stretch{1}}\large \href{mailto:jgross@mit.edu}{jgross@mit.edu} --- (631) 790-8962 --- \url{https://people.csail.mit.edu/jgross/}\hspace{\stretch{1}} \\}

  \address{\textsc{Present Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  \address{\textsc{Permanent Address}\\126 Hayrick Lane \\  Commack, NY 11725}
\end{multitex*}
\begin{multitex}[\allexceptcvlongfiles]
  \name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
      \centering Jason Gross \\
      \vspace{0.2cm}
      \large \href{https://github.com/JasonGross/}{github.com/JasonGross} \\
      \href{https://people.csail.mit.edu/jgross/}{people.csail.mit.edu/jgross} \\
      $\left.\right.$ \\
      $\left.\right.$ \\%
  }}

  \vspace{5cm}

  %\address{}%\textsc{Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  %\address{\textsc{Contact}\\\href{mailto:jgross@mit.edu}{jgross@mit.edu}\\(631) 790-8962}
  \newcommand{\addressaswidth}[1]{\makebox[\widthof{\textsc{jgross@mit.edu}}][c]{#1}}
  \address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
\end{multitex}

\begin{document}

\begin{multitex}
  \resume
\end{multitex}

\begin{multitex*}[\allexceptcvandcvlongfiles]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance of Automation in Interactive Proof Assistants, Homotopy Type Theory, Category Theory, Program Synthesis
\end{multitex*}
\begin{multitex}[curriculum-vitae]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance Engineering
\end{multitex}

\begin{multitex}
  %\section{JOB OBJECTIVE}
    %A summer position that will use my accounting and computer
    %skills.
    %
  \section{\textsc{Education}}
\end{multitex}

\begin{multitex}[\allexceptcvlongfiles]
  \employer{Massachusetts Institute of Technology}
  \dates{2013--2021}
  \title{{\rm PhD in Computer Science}}
  \location{Cambridge, MA}
  \begin{position}\\
      Advisor: Adam Chlipala\\
    \textit{Thesis: Performance Engineering of Proof-Based Software Systems at Scale}\\
    \textit{SM Thesis: An Extensible Framework for Synthesizing Efficient, Verified Parsers}
  \end{position}

  \employer{Massachusetts Institute of Technology}
  \dates{2009--2013}
  \title{{\rm BS in Mathematics and Physics}}
  \location{Cambridge, MA}
  \begin{position}\\
      GPA: 4.6/5
  \end{position}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
    \textbf{Massachusetts Institute of Technology}, Cambridge, MA  \\
    Began Doctorate of Philosophy in Computer Science in September 2013 \\
    \onlymultitex*{\cvfiles}{Research Interests: Programming Languages, Performance of Automation in Interactive Proof Assistants, Formal Verification, Homotopy Type Theory, Category Theory, Program Synthesis, Type Theory \\}
    G.P.A. 4.9/5.0 \\
    \\
    Bachelor of Science in Mathematics and Physics, June 2013   \\
    G.P.A. 4.6/5.0 \\
\end{multitex*}
\begin{multitex*}[\resumefiles]%
    \\
    Relevant Coursework: %Computer Systems
        Security, %Foundations of
        Program Analysis, Performance Engineering of Software% Systems
        , Statistical Physics, Quantum% Physics
        , Topology, Analysis, Waves \& Vibrations, Special Relativity% and the Physics of Spacetime%
        , \mbox{Algebra}% force no hyphenation
\end{multitex*}
\begin{multitex*}[\cvlongfiles]%
    \\
		Coursework:
		\begin{itemize}
			\item Computer Science: Inference and Information, Computer Systems Security, Geometric Computing, Foundations of Program Analysis, Performance Engineering of Software Systems, Structure and Interpretation of Computer Programs
			\item Mathematics: Paradox and Infinity, Category Theory for Scientists, Algebraic Topology I, Seminar in Topology, Introduction to Topology, Real Analysis, Abstract Algebra I \& II, Differential Equations, Calculus I \& II, Linear Algebra
			\item Physics: Modern Astrophysics, General Relativity (Graduate), Statistical Physics I \& II, Quantum Physics I, II, \& III, Classical Mechanics II, Waves and Vibrations, Special Relativity and the Physics of Spacetime
			\item Other Sciences: Introductory Biology, Introductory Chemistry, Design a Concentrated Solar Power Water Heater (seminar)
			\item Other: The Art and Science of Happiness, Metaphysics, Introduction to Musical Composition, Capitalism and Its Critics, Moral Psychology, Philosophy of Quantum Mechanics, Moral Problems and the Good Life, Philosophy of Love, Kaufman Teaching Certificate Program
		\end{itemize}
\end{multitex*}

% \item Presented ``Computational Higher Inductive Types: Computing with Custom Equalities'', October 6, 2013, The CSAIL Student Workshop

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Experience}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Researching program synthesis and verification with Adam Chlipala
      \item Collaboratively working on \onlymultitex*{\cvfiles}{implementing }one of the world's first algorithm-level-optimizing compilers
      \item Collaboratively implemented proven-correct crypto code now used by Google Chrome
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Experience}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Coq Development Team, INRIA}
  \dates{June 2021--Present}
  \title{Core Team Member}
  \location{Nantes, France (remote)}
  \begin{position}
    \begin{itemize}
      \item Stress Testing, Bug Reporting, Bug Minimizer, Compatibility Assurance
      \item Engineering and maintaining a bug minimizer for automatically producing minimized stand-alone test-cases from buggy code
      \item Researching performance issues that impact scalability of automated verification
      \item Have reported the plurality of all-time bugs in Coq (since 2012)
    \end{itemize}
  \end{position}

  \employer{Machine Intelligence Research Institute}
  \dates{February 2021--Present}
  \title{Research Staff}
  \location{Berkeley, CA (remote)}
  \begin{position}
    \begin{itemize}
      \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
    \end{itemize}
  \end{position}

  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{PhD Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Main Research Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
      \item Collaboratively implemented proven-correct cryptographic code now used by Google Chrome, and in the majority of secure connects from web browsers
      \item Led development of one of the world's first algorithm-level-optimizing compilers
      \item Wrote backends to C, Go, Java, and JSON; managed development of backends to Rust and Zig
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Internships}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Machine Intelligence Research Institute}
  \dates{June 2019--August 2019}
  \title{Type Theory Intern}
  \location{Berkeley, CA}
  \begin{position}
    \begin{itemize}
      \item Worked on formalizing type theories and on proving things within proof assistants
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex}
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2018--August 2018}
  \title{Software Engineering Intern}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Worked with BoringSSL on integration of proven-correct low-level ECC primitives into Chrome
    \end{itemize}
  \end{position}
  \egroup
\end{multitex}
\begin{multitex*}[\cvfiles]
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2016--September 2016}
  \title{Software Engineering Intern}
  \location{Mountain View, CA}
  \begin{position}
    \begin{itemize}
      \item Formalized low-level ECC primitives with proofs of correctness
    \end{itemize}
  \end{position}
  \egroup
\end{multitex*}
\begin{multitex}[\cvlongandmathfiles]
  \employer{MIT}
  \dates{Fall 2009--Present}
  \title{Teacher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Taught classes on \LaTeX, philosophy, linear algebra, and quantum mechanics for MIT Educational Studies Program's Splash, Spark, and Summer HSSP (High School Studies Program)
      \onlymultitex{\cvandmathfiles}{\item Teaching Assistant for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvlongandcomputersciencefiles]
  \employer{MIT}
  \dates{September 2015--December 2015}
  \title{TA for 6.172 (Performance Engineering% of Software% Systems
  )}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Created and led recitations, taught students, and helped run class
      \item Analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \employer{MIRI}
    \dates{June 12--14, 2015}
    \title{Decision Theory Workshop Attendee}
    \location{Berkeley, CA}
    \begin{position}
        \begin{itemize}
            \item Formalized various versions of L\"ob's theorem in Agda and Coq
            \item Learned about and worked on problems on the frontiers of decision theory
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex}
  \employer{Microsoft Research}
  \dates{June 2014--August 2014}
  \title{Intern}
  \location{Cambridge, United Kingdom}
  \begin{position}
    \begin{itemize}
      \item \onlymultitex*{\allexceptcvlongfiles}{Collaboratively c}\onlymultitex*{\cvlongfiles}{C}reated a language for specifying input/output behavior of x86 assembly programs\onlymultitex*{\cvlongfiles}{with Andrew Kennedy and Nick Benton}; Verified the I/O behavior of a number of simple programs
      \onlymultitex*{\cvfiles}{\item Improved automation of the x86proved library}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvfiles]
    \employer{MIT CSAIL}
    \dates{April 2012--June 2014}
    \title{Researcher}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Entered a significant amount of category theory into the automated proof assistant Coq \onlymultitex*{\cvfiles}{(\url{https://github.com/HoTT/HoTT/tree/master/theories/Categories})}
            \onlymultitex{\cvfiles}{\item Made progress towards an interface for databases and database migration on top of category theory in Coq\onlymultitex*{\cvlongfiles}{ with David Spivak and Adam Chlipala}}
            \onlymultitex*{\cvlongfiles}{\item Presented ``Building Database Management on top of Category Theory in Coq'', January 25, 2013, POPL 2013: 40th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages}
            \onlymultitex*{\cvlongfiles}{\item Presented	 ``Experience Implementing a Performant Category-Theory Library in Coq'', July 14, 2014, Interactive Theorem Proving 2014 Conference}
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{MIT CSAIL}
  \dates{November 2009--September 2011}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \onlymultitex*{\cvfiles}{\item Designed from scratch a data collection webpage, collected data for, and helped with research \onlymultitex*{\cvlongfiles}{ of Brenden Lake, Ruslan Salakhutdinov, and Josh Tenenbaum, }on categorical and transfer learning\onlymultitex*{\cvlongfiles}{ (\url{https://jgross.scripts.mit.edu/alphabets/})}}
      \onlymultitex*{\cvlongfiles,\resumefiles}{%
      \item Co-author of ``One shot learning of simple visual concepts''\onlymultitex*{\cvfiles}{ published in \emph{Proceedings of the 33rd Annual Meeting of the Cognitive Science Society}}.
        }
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT OCW}
    \dates{May--August 2010}
    \title{}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Evaluated and categorized videos of Walter Lewin solving physics problems for updating the 8.01 (Physics I) OpenCourseWare (OCW) website.
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{Commack High School}
  \dates{Fall 2006--Summer 2009}
  \title{Independent Researcher}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Independently researched circuits over sets of natural numbers for three years.
      \item Won fourth place award in mathematics in ISEF (Intel International Science and Engineering Fair) in 2009, third place award in ISEF 2008.
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \employer{Turnpike Total Appliance}
  \dates{Fall 2006--Summer 2009}
  \title{Web Page Designer}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Designed and improved main company website, took initiative to make more improvements.
    \end{itemize}
  \end{position}

\end{multitex*}

\begin{multitex}[curriculum-vitae]
\section{\textsc{Programming Languages}}
\end{multitex}
\begin{multitex*}[\allexceptcv]
\section{\textsc{Computer Skills}}
\end{multitex*}
\begin{multitex}
    \begin{itemize}
      \item
        Proficient: Coq, \onlymultitex*{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
      \item
        Working knowledge: \onlymultitex*{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
      \item
        Basic knowledge: Matlab, Lean, Idris, Ruby, Go, Ur/Web, x86 Assembly
    \end{itemize}
\end{multitex}

\begin{multitex}
  \section{\textsc{Teaching}}
  \begin{itemize}
  \item Instructor at Monsoon Math Camp: category theory, linear logic, L\"ob's theorem
  \item TA for 6.172 (Performance Engineering% of Software% Systems
    ): Led recitations, analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
  \item TA for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group
  \item Teacher at MIT ESP Programs: \LaTeX, philosophy, linear algebra, quantum mechanics
  \end{itemize}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Honors and Awards}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \begin{itemize}
      \onlymultitex*{\highschoolfiles,\cvfiles}{%
      \item
        Mathematics Honor Society (Commack High School)
      }
      \item
        Collection of 12 original K'NEX synagogue models exhibited in various museums (2004--2008)
      \onlymultitex*{\highschoolfiles}{%
        \begin{itemize}
          \item
            Collection of 12 original K'NEX synagogue models
          \item
            Volunteered to give speeches and conduct K'NEX workshops for children at opening of exhibits, as well as in many locations in NY area
          \item
            April 2008 --- Gave interview for April edition of BabagaNewz Magazine and audio interview for BabagaNewz.com
          \item
            October 2006 --- Displayed in Gurwin Nursing Home on Long Island
          \item
            February 2006 --- Exhibited at Lisa Ann Watson Children's Discovery Museum in Miami, Florida
          \item
            October 2005 --- Displayed at Shomrim (Society for New York City Jewish Police Officers)
          \item
            February--April 2005 --- Exhibited in Yeshiva University Museum in Manhattan
          \item
            December 2004 --- Displayed at Suffolk Y-JCC in Commack
        \end{itemize}
        }
      \onlymultitex*{\cvfiles}{
      \item
        Graduated Cum Laude from Commack High School (June 2009)
      }
      \onlymultitex*{\highschoolfiles}{%
      \item
        Commack High School Outstanding Academic Performance (June 2009)
      \item
        Florence Burns Scholarship for Achievement of Excellence in Computer Science (June 2009)
      \item
        Theodore R. Spedalle Award for Achievement of Excellence in Mathematics (June 2009)
      \item
        Commack High School Mathematics Research Award (June 2009)
      \item
        President's Award for Educational Excellence (June 2009)
      \item
        AP Scholar with Distinction (2009)
      \item
        Robert C. Byrd Honors Scholarship (June 2009)
      \item
        ISEF (Intel International Science \& Engineering Fair)
        \begin{itemize}
          \item
            Math research paper, ``Circuits over Sets of Natural Numbers'' won $3^\text{rd}$ Place in Math in Atlanta, May 2008
          \item
            Math research paper, ``An Investigation of the Closure of the Set of Singleton Sets of Natural Numbers under $\cup$, $\cap$, $\overline{\phantom{\emptyset}}$, $+$, $\times$'' won $4^\text{th}$ place in Math in Reno, May 2009
       \end{itemize}
      \item
        LISEF (Long Island Science \& Engineering Fair) --- Won First Place Grand Award in Math in March 2009; selected to participate in ISEF 2009
      \item
        2009 National Merit Scholarship Finalist (February 2009)
      \item
        JSHS (Junior Science \& Humanities Symposium) --- Invited to present math research paper, February 2009 and February 2008
      \item
        Rensselaer Medal --- Awarded the Rensselaer Polytechnic Institute's Rensselaer Medal, June 2008
      \item
        Outstanding Junior Award --- sole recipient, awarded by Suffolk County Math Teacher's Association (June 2008)
      \item
        County All-Star Interscholastic Math League --- competed at NYSML (New York State Mathematics League) and ARML (American Regions Mathematics League) during junior, sophomore and freshmen years, 2006--2009; Achieved individual high score for team in 2009
      \item
        Suffolk County Senior High Interscholastic Math League --- $2^\text{nd}$ in county in 2009, $1^\text{st}$ in county in 2008, $3^\text{rd}$ in county in 2007
      \item
        NYSSEF (New York State Science and Engineering Fair)
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- $1^\text{st}$ in math in Intel division, one of four first place individual winners in fair, Mu Alpha Theta award, Army award, Ohio Wesleyan University Scholarship
          \item
            2007 ($10^\text{th}$ grade) --- $1^\text{st}$ in math in Andromeda Division, earned Mu Alpha Theta Award from National Mathematics Honor Society
        \end{itemize}
      \item
        Long Island Math Fair
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- Earned gold medal for ``Group Theory'' project
          \item
            2007 ($10^\text{th}$ grade) --- Earned gold medal for ``Circuits over Sets of Natural Numbers'' project
          \item
            2006 ($9^\text{th}$ grade) --- Earned gold medal for ``Prime Numbers'' project
          \item
            2005 ($8^\text{th}$ grade) --- Earned gold medal for ``The Fourth Spatial Dimension'' project
          \item
            2004 ($7^\text{th}$ grade) --- Earned gold medal for ``The Seven Bridges of K\"onigsberg'' project
        \end{itemize}
      \item
        AIME (American Invitational Mathematics Examination)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - score of 8
          \item
            2007 ($10^\text{th}$ grade) - score of 4
          \item
            2006 ($9^\text{th}$ grade) - score of 1
        \end{itemize}
      \item
        American Mathematics Contest 12 (AMC-12)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - score of ???
          \item
            2008 ($11^\text{th}$ grade) - score of 120, top 1\% nationwide, earned title of ``School Winner''
          \item
            2007 ($10^\text{th}$ grade) - score of 103.5, top 1\% nationwide, earned title of ``School Winner''
        \end{itemize}
      \item
        American Mathematics Contest 10 (AMC-10)
        \begin{itemize}
          \item
            2006 ($9^\text{th}$ grade) - score of 127.5, top 1\% nationwide, earned title of ``School Winner''
          \item
            2005 ($8^\text{th}$ grade) - score of 118.5, earned title of ``School Winner'' in high school as an $8^\text{th}$ grader in the $9^\text{th}$ grade honors math class
        \end{itemize}
      \item
        Suffolk County Mathematics Contest
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - $1^\text{st}$ place in county competing with $12^\text{th}$ graders
          \item
            2007 ($10^\text{th}$ grade) - $4^\text{th}$ place in county competing with $11^\text{th}$ graders
          \item
            2006 ($9^\text{th}$ grade) - $4^\text{th}$ place in county competing with $10^\text{th}$ graders
          \item
            2005 ($8^\text{th}$ grade) - $5^\text{th}$ place in county competing with $9^\text{th}$ graders
          \item
            2004 ($7^\text{th}$ grade) - $1^\text{st}$ place in county
        \end{itemize}
      \item
        Suffolk County Mathematics Teachers Association (SCMTA) Tournament
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ??????th in county
          \item
            2008 ($11^\text{th}$ grade) - $4^\text{th}$ in county
          \item
            2007 ($10^\text{th}$ grade) - $2^\text{nd}$ in county
          \item
            2006 ($9^\text{th}$ grade) - $1^\text{st}$ in county
        \end{itemize}
      \item
        Commack High School Mathematics Department Awards --- June 2008, June 2007, and June 2006
      \item
        Commack Middle School Mathematics Departmental Award --- was only student of 605 students to receive Commack Middle School mathematics departmental award, 2005
      \item
        Suffolk County Mathcounts --- placed first in county as individual, 2005
      \item
        USA Mathematics Talent Search (USAMTS), 2007--2008; Placed on National List of Students with 6 Commended Solutions in 2008, silver medal in 2008
      \item
        Harvard-MIT Mathematics Tournament (HMMT), February 2009, 2008, and 2007
      \item
        New York State Science Olympiads
        \begin{itemize}
          \item
            February 2008 --- $2^\text{nd}$ place in Environmental Chemistry, $4^\text{th}$ place in Circuit Lab, $4^\text{th}$ place in Five Star Science
          \item
            February 2007 --- $4^\text{th}$ place in Circuit Lab
        \end{itemize}
      \item
        Mandelbrot Competition, team that competes in national competitions, 2006--2009; Placed on National Leaderboard (Top Tier) in 2008; co-captain in 2007--2008, captain in 2008--2009
      \item
        American Computer Science League (ACSL) Club, 2005--present; President in 2008--2009
      \item
        United States of America Computer Olympiad (USACO), 2005--present
      \item
        Suffolk County Senior High Interscholastic Math League, 2005--present
      \item
        Article was published in Teen Ink Magazine (November 2007)
      }
    \end{itemize}

\end{multitex*}

\begin{multitex*}[\highschoolfiles]
\section{\textsc{Community Service and Involvement}}
  \begin{itemize}
    \item
      Member of National Math Honor Society; Tutor students in math both in and out of school (2007--2009)
    \item
      Member of National Technology Honor Society (2007--2009)
  \end{itemize}
\end{multitex*}

\begin{multitex}
\section{\textsc{Other Activities}}
\end{multitex}
\begin{multitex}
    \begin{itemize}
      \onlymultitex*{\cvfiles,computer-science}{%
      \item Co-maintainer of the Fiat Cryptography project (\href{https://github.com/mit-plv/fiat-crypto}{\texttt{mit-plv/fiat-crypto}} on GitHub)
      \item Co-maintainer of the homotopy type theory Coq repository (\href{https://github.com/HoTT/HoTT}{\texttt{HoTT/HoTT} on GitHub})
      \item Program Committee Member of CoqPL 2022
      \onlymultitex*{\cvlongfiles}{%
        \item MIRI Decision Theory Workshop Attendee: Formalized various versions of L\"ob's theorem in Agda and Coq
        }
      \item Commiter to the SIPB BarnOwl project (\url{https://barnowl.mit.edu})
      }
      \onlymultitex*{\cvfiles}{%
      \item SIPB (Student Information and Processing Board) Member
      }
      \onlymultitex*{\cvlongfiles}{
      \item Was project leader for MITeX, an online interface for composing \LaTeX
      }
      \onlymultitex*{\cvfiles}{%
      \item President of Tech Squares, MIT's Square Dancing Club (May 2013--October 2014)
      }
      \onlymultitex*{\cvlongfiles}{%
      \item HMMT Solutions Editor (2010)
      }
      \item Canada/USA Mathcamp (Summers 2006--2009)
    \end{itemize}

\end{multitex}

\begin{multitex*}[\cvlongfiles]
% this is a hack
%\clearpage
\section{\textsc{Interests}}
    \begin{itemize}
      \item Philosophy
      \onlymultitex*{\cvlongfiles}{%
      \item Programming
      \item Dancing, especially tango, contra, and square dancing
        }
      \item Neuroscience
      \onlymultitex*{\cvlongfiles}{%
      \item Hiking
      %\onlymultitex*{\highschoolfiles}{
      \item MIT Mystery Hunt, Participated on Manic Sages Team, January 2008--2012
        %  }
        }
    \end{itemize}
\end{multitex*}

\begin{multitex}
  \nocite{*}
\end{multitex}

\begin{multitex*}[\cvshortfiles]
  \clearpage
  \section{\textsc{Selected Presentations and Publications}}
  \addtocategory{exclude}{Evaluation2021Huhmann}
  \printbibliography[title={$\left.\right.$},notcategory=exclude]
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \section{\textsc{Presentations and Publications}}
  \printbibliography[title={$\left.\right.$}]
\end{multitex*}

\begin{multitex}
\endresume
\end{multitex}
\end{document}