### 2024-11-16 misc/trash.tex
Add some generation files

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{nopageno}
\usepackage{svg}

\begin{document}
\pagestyle{empty}
% https://qci.qr-code.click/uploads/qr_codes/673826e31ca41.svg?1731733480
% https://api.qrserver.com/v1/create-qr-code/?size=150x150&data=https://jasongross.github.io/misc/trash.html

\begin{multicols}{2}
    \centering
    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
\end{multicols}

\end{document}