### 2024-11-16 misc/trash.tex
Update trash

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{nopageno}
\usepackage{svg}

\begin{document}
\pagestyle{empty}
% https://qci.qr-code.click/uploads/qr_codes/673826e31ca41.svg?1731733480
% https://api.qrserver.com/v1/create-qr-code/?size=150x150&data=https://jasongross.github.io/misc/trash.html
% https://quickchart.io/qr?text=https://jasongross.github.io/misc/trash.html&centerImageUrl=https://raw.githubusercontent.com/JasonGross/JasonGross.github.io/refs/heads/master/misc/noun-trash.png
\begin{multicols}{2}
    \centering
    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash-static.svg}
\end{multicols}

\end{document}